How environmental drivers of spatial synchrony interact

### 2023-01-10 Paper/Paper.Rnw
implemented editorial comments of Max and Jon

%***Tentative author order, Bell last for his data contributions
\author[a,*]{Daniel C. Reuman}
\author[b]{Max C.N. Castorani}
\author[c]{Kyle C. Cavanaugh}
\author[d]{Lawrence W. Sheppard}
\author[e]{Jonathan A. Walter}
\author[f,g]{Tom W. Bell}

@@

%Intro to synchrony
Spatial synchrony, the tendency for geographically disjunct populations to show correlated fluctuations 
through time, is a fundamental feature of population dynamics, linked to important topics such as 
population cycling \citep{Anderson2021}, extinction risk \citep{Ghosh2020}, and the 
stability of regional populations and ecosystem functioning \citep{Wilcox2017}. Though spatial synchrony (henceforth, 
\emph{synchrony}) has been studied for decades
in a wide variety of species ranging from viruses to plants to mammals, and at 
spatial scales ranging from centimeters to over 1000 km \citep{Liebhold2004},
recent advances in statistical methods and improvements in data availability have led to several 
major advances in our understanding of synchrony and its causes and consequences. 
For instance, the timescale structure of synchrony is now known to be 
important [\cite{Sheppard2016,Desharnais2018,Sheppard2019}, see also \cite{Keitt2008}],
synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017,Walter2020}, 
and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
is important for system stability \citep{Ghosh2020b,Ghosh2020c,Walter2021}. Our ability to infer the causes of 
synchrony is much improved in recent years \citep{Sheppard2016,Defriez2017b,Walter2017,Anderson2018,Sheppard2019}, 
and there is also growing evidence
that synchrony is changing as a newly recognized component of climate 
change \citep{Defriez2016,Sheppard2016,Hansen2020}. Spatial and community 
viewpoints on synchrony are becoming integrated, leading to a more wholistic understanding 
of the importance of synchrony for ecosystem stability \citep{Wang2014}. These new developments 
have led to an increasingly central role of the phenomenon of synchrony in many of the most 
important research topics in ecology.

%Intro to Moran effects and interactions
One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
populations. Two recent papers have documented that Moran effects of distinct 
environmental drivers can, in fact, interact, either synergistically or antagonistically,
so that total population synchrony can be either greater or less than if the distinct Moran 
drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. 
\cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
in the seas around the United Kingdom is substantially augmented by interactions between environmental 
drivers of synchrony and drivers linked with consumption by a copepod consumer.
\cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
of synchrony in giant kelp populations on the California (CA) coast, interact either synergistically 
or destructively, depending on which portion of the CA coast is examined and on 
the timescale of analysis. 
The main purpose of this study is to establish a general theory of the mechanism
of interacting Moran effects and to use examples and applications of the theory to provide 
ecological intuition for how the mechanism of interacting Moran effects works. Our goals are
distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
causes of synchrony.

@@

%Intutitive idea with the swing analogy
Having demonstrated that interactions between Moran effects can occur, we now 
use a simple analogy from everyday experience to begin to provide intuition. 
Consider $N$ children, each riding on their own playground swing and each
being pushed both by their own mother and father. 
Here, the children correspond to ecological populations and the parental pushing to environmental
influences on the populations. 
If the fathers from seperate families
were to synchronize their pushes, they would act as a Moran-like influence, tending to synchronize
the swinging motions of the children. Likewise, if the mothers from seperate families were to
synchronize their pushes, it would tend to have a separate synchronizing effect. If the pushes
of the fathers were appropriately coordinated with the pushes of the mothers, the children would swing higher, and
would also swing more synchronously, demonstrating a synergistic interaction between synchronizing
effects, i.e., a tendency for the two synchronizing effects to reinforce each other. On the 
other hand, if the maternal and paternal pushes
were not appropriately coordinated, the children 
would swing more or less randomly, with smaller amplitude, and would become asynchronous with each other. 
This second case demonstrates
an antagonistic interaction between synchronizing effects, i.e., a tendency for the two synchronizing
effects to cancel. As oscillators, children on swings are distinct in many ways from 
populations, not least because swings oscillate on one frequency/timescale, whereas populations
typically oscillate simultaneously on many timescales. 
We will see that 
a key insight is the use of a timescale-specific approach. By timescale-specific, we mean that 
fluctuations having different characteristic periods, which are superimposed in actual population time series data, 
can be understood separately. 
Timescale-specific approaches have been shown to be important to understanding synchrony and other 
phenomena (e.g., \citealp{Sheppard2016,Anderson2021,Zhao2020}). 
In the approach used in this paper, environmental and population signals are decomposed using Fourier analysis.
The simple swing analogy turns out to supply, in much idealized form, the basic intuition our formal
theory will extend.

@@

%Case study A
CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
Moran effects on different timescales, and how our theory captures that contrast.
For CaseA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:PedagogAlt},
$l_{e1}=1$q and $l_{e2}=0$q. 
The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details).
Thus, the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $1$q.
What determines the sign of interactions between Moran effects for this example is how 
this measure compares to the timescale/period, $\sigma$, being considered. 
On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
is half the period, so interaction effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
period, so $l_{e1}-l_{e2}+l_n$ is close to zero, relative to that period, and interaction effects
are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
the two noise variables reinforce each other. Comparing
Fig. \ref{fig:TheoryCaseA}a and b shows the additional influence of autoregressive population effects, which
simply multiply all Moran influences by the same timescale-dependent non-negative 
quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].

@@

%Interactions between Moran effects, CCA1
Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaseC1.
In CCA1, nitrates had $1$q-delayed positive effects on observed 
kelp populations and waves had immediate negative effects \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
just as $\epsilon^{(1)}$ effects in CaseC1 
were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
The delayed effects of nitrates in CCA1 are reflected in Table \ref{SM-tab:regcoefs}, where 
regression results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
Immediate negative effects of waves in CCA1 are also reflected in Table \ref{SM-tab:regcoefs},
where entries for $p_0^{(2)}$ are negative.
In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
tend to come in winter \citep{Schiel2015,Bell2015}. 
Thus annual nitrate peaks tend to lag annual 
wave peaks by $1$q, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
peaks by $1$q in CaseC1.
For CCA1, this is reflected in Fig. \ref{fig:noise}c, which shows that 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
has a phase of about $-\pi/2$ at the annual timescale.
Thus $l_{e1}=1$q, $l_{e2}=0$q, $l_n=1$q, and so the environmental effect alignment measures is 
$l_{e1}-l_{e2}+l_n=2$q. Lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
producing reinforcing interactions between Moran effects. 
Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
CaseC1 (compare Figs \ref{fig:TheoryCaseA}e and \ref{fig:SummaryKelpFig}c), 
and for the same reason in both cases: on long timescales, sub-annual lags
are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
effects of waves/$\epsilon^{(2)}$ therefore lead to negative interactions.
Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).

%Interactions between Moran effects, SB
Due to different nitrate effects on observable kelp growth in SB compared to 
CCA1, interactions between Moran effects followed a different 
mechanism in SB compared to CCA1, leading to the slightly negative versus positive interactions already 
documented on annual timescales for the two regions (Fig. \ref{fig:SummaryKelpFig}d v. a). 
Whereas in CCA1, nitrate effects on observed kelp density were delayed by $1$q, in SB nitrate effects 
were observable within the same quarter: 
see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. Wave effects were negative and
immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
CCA1 (Fig. \ref{fig:noise}l v. c). Thus, in SB, the environment effect alignment measure was 
$1$q, contrasting with its value of $2$q in CCA1.
Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with to the tendency of 
nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
negative wave effects which reinforced sycnhrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
while waves still had negative effects on kelp in winter. Thus the effects of nitrates and waves in SB
were approximately a quarter-cycle separated from each other with respect to the annual timescale, 
and so produced neither
much reinforcement nor much destructive interference of synchrony on 
annual timescales (Fig. \ref{fig:SummaryKelpFig}d). Slightly negative 
interactions were observed (Fig. \ref{fig:SummaryKelpFig}d) because of slight deviations from the approximate phase
alignments described above. On long timescales, interactions between Moran effects were analogous in CCA1 and SB
(though weaker in SB; Fig. \ref{fig:SummaryKelpFig}c v. f) because, again, 
sub-annual lags are inconsequential on long timescales and what mattered instead was the oppositely signed 
influences of the two environmental variables.

@@

%*DONE arguments about likelihood that interactions will be common, extending earlier arguments
We argue that it is likely that interactions between Moran effect are common. Most species
are influenced by more than one environmental driver. Drivers are frequently spatially autocorrelated,
and are also often related to each other because of their common origin in underlying climatic 
phenomena such as the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
is measured in distinct parts of the year (e.g., spring and summer temperatures, or March and April rainfall); 
or when distinct variables are measured in the same part of the 
year (e.g., spring temperature and precipitation). Future work should systematically 
investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental 
variables. Temperature and precipitation variables
measured in the same season may be particularly important candidates for interactions because of
the well known joint influence of these variables on plants.

@@

%*DONE Climate change
Climate change has the potential to influence interactions between Moran effects in two specific ways
which can be illuminated by our theory, and this potential should be investigated in
future work. Examining the third term of (\ref{rhoww_T1to3}), climate change could
alter interaction effects if it: 1) alters the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
quantifying cross synchrony between Moran drivers; or 2) alters one of the terms
$f_{\mcP^{(1)}}$ or $f_{\mcP^{(2)}}$ specifying the nature of the influence of the 
environmental variables $\epsilon^{(1)}$ and $\epsilon^{(2)}$, respectively, on populations. As advocated above,
the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ should be systematically computed in future work, for a variety 
of environmental variables, to assess whether interactions between Moran effects are likely to
be general. As part of that process, the potential for changes in $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
could also be assessed, by using either time-windowed versions or wavelet adaptations 
of this statistic, applied to either long-term climate records or future climate scanarios. 
Differences in $f_{\mcP^{(1)}}$ were responsible for differences
in the nature of Moran interactions between central CA and southern CA (Results), specifically
differences between the two regions in the lag of nitrate effects on kelp populations. 
If climate change modifies environmental effects on populations in a related way it would 
be expected to produce similarly large changes in Moran interactions. Nitrate effects on kelp 
are probably delayed by a quarter in central CA because winter waves commonly remove kelp,
and it takes time for kelp to grow to the surface and become detectable by satellite, only
then revealing the effects of elevated nitrates. In southern CA, kelp are less likely to be
completely removed by winter waves. Effects of elevated nitrates on the growth of kelp stands already reaching
the surface can be immediately observed. Climate change may similarly alter 
lags and delays associated with environmental effects on populations in at least two ways:
1) by altering species phenology; and 2) by increasing or decreasing growth rates and thereby
decreasing or increasing delays. Though it is too early to conclude that
effects on Moran interactions are among the important impacts of climate change, we feel the mechanisms
outlined above are sufficient to warrant further investigation.

>>>>

%***Tentative author order, Bell last for his data contributions
\author[a,*]{Daniel C. Reuman}
\author[b]{Max C.N. Castorani}
\author[c]{Kyle C. Cavanaugh}
\author[d]{Lawrence W. Sheppard}
\author[b,e]{Jonathan A. Walter}
\author[f,g]{Tom W. Bell}

@@

%Intro to synchrony
Spatial synchrony, the tendency for geographically disjunct populations to show correlated fluctuations 
through time, is a fundamental feature of population dynamics, linked to important topics such as 
population cycling \citep{Anderson2021}, extinction risk \citep{Ghosh2020}, and the 
stability of regional populations and ecosystem functioning \citep{Wilcox2017}. Though spatial synchrony (henceforth, 
\emph{synchrony}) has been studied for decades
in a wide variety of species ranging from viruses and plants to mammals, and at 
spatial scales ranging from centimeters to over 1000 km \citep{Liebhold2004},
recent advances in statistical methods and improvements in data availability have led to several 
major advances in our understanding of synchrony and its causes and consequences. 
For instance, the timescale structure of synchrony is now known to be 
important [\cite{Sheppard2016,Desharnais2018,Sheppard2019}, see also \cite{Keitt2008}],
synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017,Walter2020}, 
and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
is important for system stability \citep{Ghosh2020b,Ghosh2020c,Walter2021}. Our ability to infer the causes of 
synchrony is much improved in recent years \citep{Sheppard2016,Defriez2017b,Walter2017,Anderson2018,Sheppard2019}, 
and there is also growing evidence
that synchrony is changing as a newly recognized component of climate 
change \citep{Defriez2016,Sheppard2016,Hansen2020}. 
Distinct viewpoints on synchrony from population and community ecology are 
becoming integrated, leading to a more wholistic understanding 
of the importance of synchrony for ecosystem stability \citep{Wang2014}. These new developments 
have led to an increasingly central role of the phenomenon of synchrony in many of the most 
important research topics in ecology.

%Intro to Moran effects and interactions
One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
populations. Two recent papers have documented that Moran effects of distinct 
environmental drivers can, in fact, interact, either synergistically or antagonistically,
so that total population synchrony can be either greater than or less than what synchrony would be if the distinct Moran 
drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. 
\cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
in the seas around the United Kingdom is substantially augmented by interactions between environmental 
drivers of synchrony and drivers linked with consumption by a copepod consumer.
\cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
of synchrony in giant kelp populations on the California (CA) coast, interact either synergistically 
or destructively, depending on which portion of the CA coast is examined and on 
the timescale of analysis. 
The main purpose of this study is to establish a general theory of the mechanism
of interacting Moran effects and to use examples and applications of the theory to provide 
ecological intuition for how the mechanism of interacting Moran effects works. Our goals are
distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
causes of synchrony.

@@

%Intutitive idea with the swing analogy
Having demonstrated that interactions between Moran effects can occur, we now 
use a simple analogy from common experience to begin to provide intuition. 
Consider $N$ children, each riding on their own playground swing and each
being pushed both by their own mother and their own father. 
Here, the children correspond to ecological populations and the parental pushing to environmental
influences on the populations. 
If the fathers from seperate families
were to synchronize their pushes, they would act as a Moran-like influence, tending to synchronize
the swinging motions of the children. Likewise, if the mothers from seperate families were to
synchronize their pushes, it would tend to have a separate synchronizing effect. If the pushes
of the fathers were appropriately coordinated with the pushes of the mothers (either happening at the same
time, if the mothers and fathers are standing on the same sides of the swining children, or at
opposite times in the swing period if the parents are on opposite sides), the children would swing higher, and
would also swing more synchronously, demonstrating a synergistic interaction between synchronizing
effects, i.e., a tendency for the two synchronizing effects to reinforce each other. On the 
other hand, if the maternal and paternal pushes
were not appropriately coordinated, the children 
would swing more or less randomly, with smaller amplitude, and would become asynchronous with each other. 
This second case demonstrates
an antagonistic interaction between synchronizing effects, i.e., a tendency for the two synchronizing
effects to cancel. As oscillators, children on swings are distinct in many ways from 
populations, not least because swings oscillate on one frequency/timescale, whereas populations
typically oscillate simultaneously on many timescales. 
We will see that 
a key insight is the use of a timescale-specific approach. By timescale-specific, we mean that 
fluctuations having different characteristic periods, which are superimposed in actual population time series data, 
can be understood separately. 
Timescale-specific approaches have been shown to be important to understanding synchrony and other 
phenomena (e.g., \citealp{Sheppard2016,Anderson2021,Zhao2020}). 
In the approach used in this paper, environmental and population signals are decomposed using Fourier analysis.
The simple swing analogy turns out to supply, in much idealized form, the basic intuition our formal
theory will extend.

@@

%Case study A
CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
Moran effects on different timescales, and how our theory captures that contrast.
For CaseA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:PedagogAlt},
$l_{e1}=1$q and $l_{e2}=0$q. 
The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details).
Thus, the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $1$q.
What determines the sign of interactions between Moran effects for this example is how 
this measure compares to the timescale/period, $\sigma$, being considered. 
On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
is half the period, so interaction effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
period, so $l_{e1}-l_{e2}+l_n$ is close to zero, relative to that period, and interaction effects
are positive: relative to long timescales, $\epsilon^{(1)}$ 
and $\epsilon^{(2)}$ effects happen close to simultaneously, so
the two noise variables reinforce each other. Comparing
Fig. \ref{fig:TheoryCaseA}a and b shows the additional influence of autoregressive population effects, which
simply multiply all Moran influences by the same timescale-dependent non-negative 
quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].

@@

%Interactions between Moran effects, CCA1
Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaseC1.
In CCA1, nitrates had $1$q-delayed positive effects on observed 
kelp populations and waves had immediate negative effects \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
just as $\epsilon^{(1)}$ effects in CaseC1 
were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
The delayed effects of nitrates in CCA1 are reflected in Table \ref{SM-tab:regcoefs}, where 
regression results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
Immediate negative effects of waves in CCA1 are also reflected in Table \ref{SM-tab:regcoefs},
where entries for $p_0^{(2)}$ are negative.
In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
tend to come in winter \citep{Schiel2015,Bell2015}. 
Thus annual nitrate peaks tend to lag annual 
wave peaks by $1$q, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
peaks by $1$q in CaseC1.
For CCA1, this is reflected in Fig. \ref{fig:noise}c, which shows that 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
has a phase of about $-\pi/2$ at the annual timescale.
Thus $l_{e1}=1$q, $l_{e2}=0$q, $l_n=1$q, and so the environmental effect alignment measure is 
$l_{e1}-l_{e2}+l_n=2$q. Lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
producing reinforcing interactions between Moran effects. 
Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
CaseC1 (compare Figs \ref{fig:TheoryCaseA}e and \ref{fig:SummaryKelpFig}c), 
and for the same reason in both cases: on long timescales, sub-annual lags
are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
effects of waves/$\epsilon^{(2)}$ therefore lead to negative interactions.
Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).

%Interactions between Moran effects, SB
Due to different nitrate effects on observable kelp growth in SB compared to 
CCA1, interactions between Moran effects followed a different 
mechanism in SB compared to CCA1, leading to the slightly negative versus positive interactions already 
documented on annual timescales for the two regions (Fig. \ref{fig:SummaryKelpFig}d v. a). 
Whereas in CCA1, nitrate effects on observed kelp density were delayed by $1$q, in SB nitrate effects 
were observable within the same quarter: 
see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. Wave effects were negative and
immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
CCA1 (Fig. \ref{fig:noise}l v. c). Thus, in SB, the environment effect alignment measure was 
$1$q, contrasting with its value of $2$q in CCA1.
Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with the tendency of 
nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
negative wave effects which reinforced sycnhrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
while waves still had negative effects on kelp in winter. Thus the effects of nitrates and waves in SB
were approximately a quarter-cycle separated from each other with respect to the annual timescale, 
and so produced neither
much reinforcement nor much destructive interference of synchrony on 
annual timescales (Fig. \ref{fig:SummaryKelpFig}d). Slightly negative 
interactions were observed (Fig. \ref{fig:SummaryKelpFig}d) because of slight deviations from the approximate phase
alignments described above. On long timescales, interactions between Moran effects were analogous in CCA1 and SB
(though weaker in SB; Fig. \ref{fig:SummaryKelpFig}c v. f) because, again, 
sub-annual lags are inconsequential on long timescales and what mattered instead was the oppositely signed 
influences of the two environmental variables.

@@

%*DONE arguments about likelihood that interactions will be common, extending earlier arguments
We argue that it is likely that interactions between Moran effect are common. Most species
are influenced by more than one environmental driver. Drivers are frequently spatially autocorrelated,
and are also often related to each other because of their common origin in underlying climatic 
phenomena such as the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
is measured in distinct parts of the year (e.g., spring and summer temperatures, or March and April rainfall); 
or when distinct variables are measured in the same part of the 
year (e.g., spring temperature and precipitation). Such scenarios involving seasonality of effects, 
which were recently explored by \cite{WalterInPress}, can produce specific manifestations of the general mechanisms
explored here. \cite{WalterInPress} found interactions and cross synchrony to be important.
Future work should systematically 
investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental 
variables. Temperature and precipitation variables
measured in the same season may be particularly important candidates for interactions because of
the well known joint influence of these variables on plants.

@@

%*DONE Climate change
Climate change has the potential to influence interactions between Moran effects in two specific ways
which can be illuminated by our theory, and this potential should be investigated in
future work. Examining the third term of (\ref{rhoww_T1to3}), climate change could
alter interaction effects if it: 1) alters the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
quantifying cross synchrony between Moran drivers; or 2) alters one of the terms
$f_{\mcP^{(1)}}$ or $f_{\mcP^{(2)}}$ specifying the nature of the influence of the 
environmental variables $\epsilon^{(1)}$ and $\epsilon^{(2)}$, respectively, on populations. As advocated above,
the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ should be systematically computed in future work, for a variety 
of environmental variables, to assess whether interactions between Moran effects are likely to
be general. As part of that process, the potential for changes in $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
could also be assessed, by using either time-windowed versions or wavelet adaptations 
of this statistic, applied to either long-term climate records or future climate scanarios. 
Differences in $f_{\mcP^{(1)}}$ were responsible for differences
in the nature of Moran interactions between central CA and southern CA (Results), specifically
differences between the two regions in the lag of nitrate effects on kelp populations. 
If climate change modifies environmental effects on populations in a related way it would 
be expected to produce similarly large changes in Moran interactions. Nitrate effects on kelp 
are probably delayed by a quarter in central CA because winter waves commonly remove kelp,
and it takes time for kelp to grow to the surface and become detectable by satellite, only
then revealing the effects of elevated nitrates \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
In southern CA, kelp are less likely to be completely removed by winter waves. 
Effects of elevated nitrates on the growth of kelp stands already reaching
the surface can be immediately observed \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
Climate change may similarly alter 
lags and delays associated with environmental effects on populations in at least two ways:
1) by altering species phenology; and 2) by increasing or decreasing growth rates and thereby
decreasing or increasing delays. Though it is too early to conclude that
effects on Moran interactions are among the important impacts of climate change, we feel the mechanisms
outlined above are sufficient to warrant further investigation.
